---
title: "First Day Notes"
author: "Patrick Kimes"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    df_print: paged
---

# Outline

Here we try out doing an analysis.

```{r}
library(tidyverse)
library(lubridate)
library(cowplot)
library(bigrquery)
library(ggbeeswarm)

theme_set(theme_minimal())
theme_x_rotate <- function(...) {
    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1), ...)
}

## bigrquery issue: https://github.com/r-dbi/bigrquery/issues/395
options(scipen = 20)
```

We will be querying from the same project and dataset on BigQuery.

```{r}
project <- "hst-953-2019"
bq_auth(email = "patrick.kimes@gmail.com")
```

Data tables will be saved to a local directory for easier future access.

```{r}
bqdir <- file.path("..", "data", "bigquery")
dir.create(bqdir, showWarnings = FALSE, recursive = TRUE)
```

# Data

## Cohort

We don't care about notes, so we'll flatten this table and also obtain the
length of stay for each admission.

```{r}
firstday_rds <- file.path(bqdir, "firstday.rds")
if (file.exists(firstday_rds)) {
    firstday <- readRDS(firstday_rds)
} else {
    query <- read_file(file.path("..", "sql", "20210417-firstnotes.sql"))
    res <- bq_project_query(x = "hst-953-2019", query = query)
    firstday <- bq_table_download(res, page_size = 1e3)
    saveRDS(firstday, firstday_rds)
}
```

Also filter out all visits without any first day notes.

```{r}
firstdaye_rds <- file.path(bqdir, "firstdaye.rds")
if (file.exists(firstdaye_rds)) {
    firstdaye <- readRDS(firstdaye_rds)
} else {
    firstdaye <- firstday %>%
        dplyr::filter(!is.na(CHARTDATE))
    saveRDS(firstdaye, firstdaye_rds)
}
```

# Plots

Count up number of notes.

```{r, fig.width = 6, fig.height = 3.5}
firstday %>%
    dplyr::group_by(SUBJECT_ID, HADM_ID) %>%
    dplyr::summarize(nnotes = sum(!is.na(CHARTDATE))) %>%
    dplyr::ungroup() %>%
    dplyr::count(nnotes, name = "nadmit") %>% 
    dplyr::arrange(nnotes)
    
firstday %>%
    dplyr::group_by(SUBJECT_ID, HADM_ID) %>%
    dplyr::summarize(n = sum(!is.na(CHARTDATE))) %>%
    dplyr::ungroup() %>% 
    ggplot(aes(x = n)) +
    geom_histogram(boundary = 1/2, binwidth = 1,
                   color = 'white', fill = 'gray20') +
    xlab("number of notes") +
    ylab("number of admissions") + 
    ggtitle("First day notes per admission")

firstdaye %>%
    dplyr::count(SUBJECT_ID, HADM_ID) %>%
    ggplot(aes(x = n)) +
    geom_histogram(boundary = 1/2, binwidth = 1,
                   color = 'white', fill = 'gray20') +
    xlab("number of notes") +
    ylab("number of admissions") + 
    ggtitle("First day notes")
```

Check age.

```{r}
firstdaye %>%
    dplyr::count(AVG_OASIS_AGE >= 18)
```

