---
title: "First Day Notes"
author: "Patrick Kimes"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    df_print: paged
---

# Outline

Here we try out doing an analysis.

```{r}
library(tidyverse)
library(bigrquery)

theme_set(theme_minimal())
theme_x_rotate <- function(...) {
    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1), ...)
}

## bigrquery issue: https://github.com/r-dbi/bigrquery/issues/395
options(scipen = 20)
```

We will be querying from the same project and dataset on BigQuery.

```{r}
project <- "hst-953-2019"
bq_auth(email = "patrick.kimes@gmail.com")
```

Data tables will be saved to a local directory for easier future access.

```{r}
bqdir <- file.path("..", "data", "bigquery")
dir.create(bqdir, showWarnings = FALSE, recursive = TRUE)
```

# Data

Need time of ICU stay in/out to match with notes.

```{r}
query <- "
SELECT SUBJECT_ID, HADM_ID, CHARTTIME, CATEGORY
FROM `physionet-data.mimiciii_notes.noteevents`
"
res <- bq_project_query(x = "hst-953-2019", query = query)
notetimes <- bq_table_download(res, page_size = 1e4)
```

Need ICU in/out times.

```{r}
query <- "
SELECT ICUSTAY_ID, HADM_ID, INTIME, OUTTIME
FROM `physionet-data.mimiciii_clinical.icustays`
"
res <- bq_project_query(x = "hst-953-2019", query = query)
icutimes <- bq_table_download(res, page_size = 1e4)
```

Read in table of IDs that people are asking about.

```{r}
intab <- read_csv(file.path("..", "data", "external", "transfusion_ids.csv"),
                  col_types = cols())
```

Only use some columns of their table.

```{r}
intab <- intab %>%
    dplyr::select(-transfusedmarker,
                  -charttime, -chartdate) %>%
    dplyr::rename_all(toupper)
```

Add ICU in/out times.

```{r}
intab <- intab %>%
    dplyr::left_join(icutimes, by = c("ICUSTAY_ID", "HADM_ID"))
```

## HADM_ID level compare

Look for notes over these windows.

```{r}
noteset <- notetimes %>%
    dplyr::left_join(intab, by = c("HADM_ID")) %>%
    dplyr::filter(!is.na(CHARTTIME)) %>%
    dplyr::filter(!is.na(HAS.NOTES)) %>%
    dplyr::filter(dplyr::between(CHARTTIME, INTIME, OUTTIME))

noteset %>%
    dplyr::count(HADM_ID, ICUSTAY_ID)
```

Not many.

# HADM_ID level compare

This table is by ICU stay. Notes are by subject ID. Just summarize.

```{r}
intab %>%
    dplyr::group_by(HADM_ID) %>%
    dplyr::filter(n() > 1L)
```

Interestingly, hospital admission IDs are unique.. So we'll just ignore
the ICU stay IDs and take the columns we care about.

```{r}
intab <- intab %>%
    dplyr::mutate(has.notes = as.logical(has.notes),
                  hadm_id = as.numeric(hadm_id)) %>%
    dplyr::select(HADM_ID = hadm_id, has.notes)
```

Now check if really have no notes.

```{r}
comparetab <- intab %>%
    dplyr::select(HADM_ID, HAS.NOTES) %>%
    dplyr::left_join(nnotes, by = "HADM_ID") %>%
    dplyr::mutate(N = TRUE) %>%
    tidyr::complete(nesting(HADM_ID, SUBJECT_ID, HAS.NOTES),
                    CATEGORY,
                    fill = list(N = FALSE)) %>%
    dplyr::filter(!is.na(CATEGORY)) 

comparetab %>%
    dplyr::group_by(CATEGORY) %>%
    dplyr::summarize(agree = mean(HAS.NOTES == N))
```

Even at the HADM_ID level, no agreement.

```{r}
comparetab %>%
    dplyr::group_by(HADM_ID, SUBJECT_ID, HAS.NOTES) %>%
    dplyr::summarize(ACTUAL.NOTES = any(N)) %>%
    dplyr::ungroup() %>%
    dplyr::count(HAS.NOTES, ACTUAL.NOTES)
```

Again, not the right agreement.

Check other class of notes.

```{r}
comparetab %>%
    dplyr::filter(CATEGORY %in% c("Physician ", "Nursing",
                                  "Discharge summary",
                                  "Nursing/other")) %>%
    dplyr::group_by(HADM_ID, SUBJECT_ID, HAS.NOTES) %>%
    dplyr::summarize(ACTUAL.NOTES = any(N)) %>%
    dplyr::ungroup() %>%
    dplyr::count(HAS.NOTES, ACTUAL.NOTES)
```
